### 2014-12-09 tex/short_2015/body.tex
Now 4 pages

\begin{abstract}
\color{red}
   We describe an algorithm and implementation of automatic proxy model
   generation for a client/server remote rendering setup. A {\em proxy model} is
   a light-weight version of a main 3D model that is cheaper to render. We
   automatically derive this from the main model. The client is typically a
   tablet, for which we only assume availability of WebGL and Javascript. The
   server makes use of OpenGL and a web server. When the rate of received
   server-rendered images deteriorate, the client renders the proxy model, which
   is computed from depth buffers bundled with rendered images from the
   server. This algorithm requires little modification of the application
   itself.
   % Husk: Skal vaere en blank linje nedenfor

@@

With increasingly powerful rendering engines, in forms ranging from mobile and
hand-held devices to traditional stationary computers, one might be led to
believe that rendering frame rates could only increase and visualization could
only become more interactive. Alas, hand in hand with this development we also
find ever increasing requirements on computational accuracy, power efficiency,
data sizes, scaling, etc.

One effect of this, also enhanced by new cloud-based approaches, wireless
connection, and mobile devices with display capabilities outpacing their
computation and storage capabilities, is that interactivity still is a difficult
issue. In this work, we consider a client/server model for rendering, but this
is just a way of separating the interaction from the rendering; the discussion
also applies to a more integrated setup.  In order to obtain the frame rates on
the client suitable for interactive visualizations, demands are put on
bandwidth, latency and rendering speed. If one or more of these demands are not
met, according to a suitable measure, the frame rate will deteriorate together
with the user experience.

One way to deal with such situations is for the client to have a {\em proxy
model}, defined as a temporary model to be shown and manipulated locally on the
client while waiting for the appropriate image from the server.  A particularly
unwelcomed chore is that of building proxy geometries for various data sets. The
typical ``hello world'' proxy geometry is the familiar bounding box wire frame,
but this will degrade the user experience unacceptably for most
applications. One alternative is to hand craft such simpler proxies. They have
to be simple both for the building cost, and also for the actual rendering on
the possibly very thin client. Another alternative is to automate this process,
a large triangulation may for instance be represented by a much coarser
approximating triangulation that can be built automatically. One disadvantage of
this approach is that this is not easy to do for many kinds of data, like our
main case in which we have an oil reservoir viewer (FRView,~\cite{cloudviz})
that renders large {\em corner point grids}, together with faults, oil wells,
and more. See Figure~\ref{fig:FRView} above for an example of both a fully
completed rendering, and automatically generated proxy geometries rendered in
Google Chrome.

Our suggested solution is to pass depth information from the server along with
ordinary rendered images. From this depth image, a rudimentary 3D model can be
built, and using the ordinary image as a texture, this model can be transformed
and rendered on the client while waiting for the next update from the server. If
the server does not take too long to respond, and/or the client does not change
the position or orientation of the model too much, this display of a proxy model
integrates seemlessly with the slower stream of server-rendered frames.

We also mention that even if bandwidth and latency is not a problem, it may be
desirable to let a server of limited capacity serve many simultaneous users,
hence decreasing the effective server time available for each one. In this case,
suitable scaling properties may still be achieved without increasing server
capacity. The results described in this paper will be commercialized in the new
SINTEF project \mbox{ArrowHead}~\cite{arrowhead}.

@@

Our approach has some similarities to {\em image based rendering} (IBR)
techniques, with the difference that an important IBR problem would be the
reconstruction of a depth map from images, while we have access to the full
depth map from a rendering pass.  Another way to use depth maps similar to what
we do, is for ``immersive streaming'', see \eg,~\cite{ibr}, where focus is on
depth map compression, an issue that we also consider. {\color{red}Another work
focused on similar streaming and geometry compression,
is~\cite{teler}. Also,~\cite{220764} contains some of the same ideas as our
work, with respect to image-based rendering accelleration.}

Common to many IBR-algorithms is also that of stitching together 3D or 2.5D
point clouds. We could also in principle do this for our 2.5D maps fetched from
the server, but it is not clear that the benefit would outweight the
cost.

We are not aware of anybody else using depth maps exactly for our purpose, most
research seems to be focused on the retrieval of the depth map from images,
while we use an existing depth map to distill and render temporary geometry.

{\color{red}Finally, we note that these 2.5D heightmaps are exactly what ``3D cameras''
(time-of-flight and other range image sensors) produce, but most authors
considering these are typically building more complex geometries before
visualization.} % , see~\eg,~\cite{}
\marginpar{\textbf{REF?!}}

@@

The main idea is to automatically construct a proxy geometry on the client side
from depth data provided by the server. The server already produces a depth
buffer as part of its rendering process.  This means that our technique is not
immediately ready for \eg, ray-tracing.) This depth buffer may be regarded as a
height map over a viewing plane relative to the observing client. Hence, this
depth map does not contain information about scene elements not in view for the
set of view parameters at the time of server side rendering. Our approach is
based on the assumption that small transformations of this height map still will
give good approximations to the scene being rendered. These renderings are
exactly what the client will compute and show while waiting for the next update
from the server. See Figure~\ref{fig:2Dheightmap} and
Figure~\ref{fig:2DheightmapRotated} below for an illustration of this heightmap
computed on the server, and a recovered 3D model on the client.

\begin{figure}[htb]
  \centering
  \input{fig2.tex}
  \caption{\label{fig:2Dheightmap} The rendered image (of \eg, a cube) as a
           height map seen from above, for a certain transformation. Also shown
           are the corresponding RGB-fragments (red and green) and depth
           fragments (floating point scalars shown as grayscale pixels.)}
\end{figure}

In Figure~\ref{fig:2DheightmapRotated} below, a sequence of three server-rendered
images (solid lines, red and blue) is shown, together with intermediate
client-rendered proxy models with different features. These features will be
discussed later, in Section~\ref{sec:client}.

\begin{figure}[htb]
  \centering
  \subfigure[Low frame rate]{
    \input{fig3a.tex}
  }
  \subfigure[Initial proxy model]{
    \input{fig3b.tex}
  }
  \subfigure[Client-generated frame]{
    \input{fig3c.tex}
  }
  \subfigure[With texturing]{
    \input{fig3d.tex}
  }
  \subfigure[Screen-space-sized splats]{
    \input{fig3e.tex}
  }
  \subfigure[With intra-splat depths]{
    \input{fig3f.tex}
  }
  \caption{\label{fig:2DheightmapRotated}
           The depth map and image sent from the server enables the client to
           recover a 3D model approximating the one rendered on the server. This
           client-side model is shown here with a small additional rotation, again
           seen from above.}
\end{figure}

Note that the depth map from the server does not allow the client to recover all
information in the server side model. For instance, the solid surface shown in
Figure~\ref{fig:2Dheightmap} (\ie, the {\em topology}) cannot be deduced, hence
we show the dots in Figure~\ref{fig:2DheightmapRotated}~(b). Still, there are
several things we can do to improve the client's approximation of
the server-side image, and these will now be described.

%-------------------------------------------------------------------------
\subsection{The server}

The server keeps a 3D model and accompanying viewing parameters, and will render
an image for delivery to the client either on its own accord or because the
client requests it. As long as the server renders a picture into a framebuffer
with a connected depth buffer, it has a negligible cost for the server to bundle
the resulting depth map with the rendered image upon return of the result. In
practice, this is easily handled by a few extra lines of computer code where the
image is extracted from its framebuffer and prepared for sending to the
client.

Since the depth buffer adds somewhat to the data being sent, it is important to
keep it to a minimum. We have found that reducing the spatial resolution of the
depth buffer (for instance by a factor of $1/16$) only to a negligible extent
degrades the proxy model being reconstructed by the client. We encode each depth
value in the range $[0, 1]$ as a 16 bit fixed point number, this also works
quite well. These measures cause the bundling of the depth buffer with the rgb
image to impose just a small data overhead.

Further compression may bring this data size down even more, but the cost of
compression/decompression must also be added to the cost of rendering on the
client/server. In~\cite{DBLP:journals/tvcg/Lindstrom14} a suitable algorithm is
proposed that promises to be fast both for the compression and decompression
stages. \marginpar{\textbf{REF?!}} Together with the rendered rgb image and the
depth image we will also send the current view transformations back to the
client.

%-------------------------------------------------------------------------
\subsection{The client}
\label{sec:client}

When the client receieves an rgb and depth image, together with view
transformations, it builds a ``proxy model'' from this. This model can then be
transformed and rendered directly if there is no available image from the
server, or it can be combined with other proxy models the client already has in
store. The latter option will be discussed in
Section~\ref{sec:proxyModelReplacement}.

%-------------------------------------------------------------------------
\subsubsection{Proxy model building, splats}

\marginpar{\textbf{boer kunne krympes}}
As indicated by Figure~\ref{fig:2DheightmapRotated}~(b), the received
height map does not really allow us more than concluding where a set of points
belong to the 3D model, these points giving rise to exactly the depth map from
the server. Given these points, the client may build a model approximating the
actual scene being rendered on the server. Since the information from such a
depth map can only contain the foremost point along any ray from the observer
toward the scene, it is often said to be in 2.5D, as opposed to purely 2D or
fully 3D.

The most simple thing the client can do, is just to transform and render this
set of 3D points with the color sampled from the corresponding location in the
server-rendered image, this is illustrated in
Figure~\ref{fig:2DheightmapRotated}~(b).

It makes sense for the client only to render at most one 3D point for each depth
fragment available. This may be too much for a thin client, and one may want to
operate with a smaller number of 3D points, and instead render each of them with
a larger number of pixels on the client side, this is called a {\em splat}. A
set of such splats for a given depth image is what we have termed a {\em proxy
model}. There are some issues to discuss in relation to this, \eg, what size and
shape of geometries to render for each 3D point, how to color them, how to
compute depth fragments and so on.  By rendering splats, we get fewer ``false
connections'' than if we connect 3D points and reconstruct topology, for
instance in the shape of a triangulation. On the other hand, we risk getting
more ``holes'' in our models.

The simplest solution is to render each 3D point as a fixed geometry, for
instance a 2D disk or rectangle, with the correspondig color from the image, see
Figure~\ref{fig:2DheightmapRotated}~(c), where we have added the transformation
local to the client, which the server has not been able to provide an image for.
We will briefly describe some improvements to this. We also note that other
possibilities exist apart from this ``splat-based'' proxy model, \eg, building
and maintaining a 3D occupancy mesh, computing a distance field from which
iso-surfaces can be extracted, etc.

@@

Our splats may result in many client window fragments, necessitating an
``intra-splat'' fragment coloring or texturing. This is not entirely straight
forward, since the client really do not know how a finitely-sized splat should
map to the server's geometry, having only a discrete set of 3D points on that
model.

A first approximation is for the client to assume that the corresponding part of
the server's model is planar in a region around the given point. If this is the
case, a local 2D texture transformation will provide a good approximation to the
intra-splat texturing to be performed on the client. Let $P_c$ and $P_s$ be
projection matrices on the client and server, respectively, and $M_c$ and $M_s$
corresponding view matrices. Let there be $n_x \times n_y$ splats laid out in a
grid. For the splat centered in $\vv_{i, j} = (x_j, y_i, z_{i, j})$, to be
centered on the client's canvas at screen coordinate $\pv_{i, j} = P_c M_c
M_s^{-1} P_s^{-1}\vv_{i, j} = U\vv_{i, j}$, the texture coordinate
transformation to be used is,
\[
  T =
  \begin{pmatrix}
    \frac{1}{n_x} & 0 \\
    0 & \frac{1}{n_y}
  \end{pmatrix}
  \Big( \sv_x \, \, \, \sv_y \Big)^{-1}
  \begin{pmatrix} 
    \frac{\text{vpwidth}}{n_x} & 0 \\
     0 & \frac{\text{vpheight}}{n_y}
  \end{pmatrix}
  \cdot \text{overlap},
\]
\[
   =
  A
  \Big( \sv_x \, \, \, \sv_y \Big)^{-1}
  B
  \cdot \text{overlap},
\]
where $A$ maps the ``client's splat region'' (in $[0, 1]^2$) to the
corresponding texture area, $(\sv_{x} \, \, \, \sv_{y})^{-1}$ maps the
``client's screen space splat area'' to $[0, 1]^2$, and $B$ together with the
``overlap'' just provides a scaling factor so that ``overlap=1'' produces
non-overlapping but covering splats when $M_c=M_s$, \ie, the client does no
transformation of the proxy model.
%
% After mult by proj matrix: clip coo, range of (x, y, z) = [-w, w]^3
% After persp division: ndc [-1, 1]^3
% After viewport transf: window coordinates
% (Range [vp[0] vp[1]] x [[vp[2] vp[3]] x [0, 1] ?)
%
We compute $\sv_{x}$ and $\sv_{y}$ by evaluating proxy model positions and
performing the perspective division and transformation into window coordinates
in the vertex shader. Let proxy model positions corresponding to three corners
of a quadrilateral splat in clip coordinates be
given by
\[
  \pv =
  U { \small \begin{pmatrix} x \\ y \\ d \\ 1 \end{pmatrix} },
  \pv_{x+\Delta x} =
  U { \small \begin{pmatrix} x+\Delta x \\ y \\ d_{\Delta x} \\ 1 \end{pmatrix} }
  \text{and}\, \, \, 
  \pv_{y+\Delta y} =
  U { \small \begin{pmatrix} x \\ y+\Delta y \\ d_{\Delta y} \\ 1 \end{pmatrix} },
\]
where $d = 2\text{depthTex}(x, y) - 1$, $d_{\Delta x} =
2\text{depthTex}(x+\Delta x, y) - 1$, and $d_{\Delta y} = 2\text{depthTex}(x,
y+\Delta y) - 1$ are depths sampled from the received depth buffer
$\text{depthTex}$ and transformed to $[-1, 1]$. By combining this with a
$w$-component set to one, we have in effect done a perspective division, so that
the right-hand side vertices are in clip space, suitable for multiplication with
$U$.

We use
\begin{equation}
  \begin{aligned}
    \Delta x & = n_x / \text{width}(\text{depthTex}) \\
    \Delta y & = n_y / \text{height}(\text{depthTex}),
  \end{aligned}
  \label{eq:deltaChoice}
\end{equation}
note that this choice of $\Delta x$ and $\Delta y$ is not unique, we might have
chosen something larger. The values are used for looking up depth image samples,
so we should not choose something too small, our choice is exactly to look for
neighbouring depth samples. Since this calculation of $\sv_x$ and $\sv_y$ is
really a difference approximating a gradient on the geometry, we should not make
$\Delta x$ and $\Delta y$ to large either. Experiments show us
that~(\ref{eq:deltaChoice}) is really a good choice.

This then leaves us $\pv$, $\pv_{x+\Delta x}$ and $\pv_{y+\Delta y}$ in clip
coordinates, and after perspective division and viewport transformation, we get
corresponding window coordinates $\sv$, $\sv_{x+\Delta x}$ and $\sv_{y+\Delta
y}$, which we can subtract to obtain 2D vectors spanning the splat,
\[
  \sv_{\delta} =
  \sv_{x+\delta} - \sv =
    \frac{L}{2\delta} \left(
        \frac{\pv_{x+\delta}.xy}{\pv_{x+\delta}.w} -
        \frac{\pv.xy}{\pv.w}
    \right),
\]
where $\delta$ is either $\Delta x$ or $\Delta y$, $L$ is either
$\text{width}$ or $\text{height}$, and we have used the ``shader
notation'' to extract 2D vectors and scalars from the 4D vectors, and
$\text{width}$ and $\text{height}$ are the viewport width and height,
respectively.

Then, on the client we render a \texttt{glPoint} for each splat, with texture
coordinates $(s, t)^\prime$ and \texttt{glPointSize} set to a larger value that will
produce many fragments. Each fragment will then look up the server-rendered
image at position
\[
  \begin{pmatrix}
    s \\ t
  \end{pmatrix} +
  T 
  \begin{pmatrix}
    u \\ v
  \end{pmatrix}
  =
  \begin{pmatrix}
    s \\ t
  \end{pmatrix} +
  T 
  \begin{pmatrix}
    \texttt{glPointCoord}-\frac{1}{2}
  \end{pmatrix}
\]
where $(u, v)$ are ``intra-splat texture coordinates''.

When the assumption that the geometry is locally planar does not hold, \eg, if
the splats are very large, or they originate from a rather curved or non-smooth
part of the geometry, this may look rather odd, see
Figure~\ref{fig:LargeSplatsOnCorners}.

\begin{figure}[htb]
  \centering
  \subfigure[Server-rendered]{
    \includegraphics[width=.3\linewidth]{splat1.jpg}
  }
  \subfigure[$P_c M_c = P_s M_s$]{
    \includegraphics[width=.3\linewidth]{splat2.jpg}
  }
  \subfigure[$P_c M_c \neq P_s M_s$]{
    \includegraphics[width=.3\linewidth]{splat3.jpg}
  }
  \caption{\label{fig:LargeSplatsOnCorners}
           Splats for which depths are sampled on a planar region look nice on
  exactly this region, but contorted elsewhere, notice the corner and edges of
  the cube.}
\end{figure}

The problem illustrated in Figure~\ref{fig:LargeSplatsOnCorners}, is that the
geometry is far from planar, and our simple texture coordinate transformation
will produce the wrong result for the corresponding part of involved
splats. Some ways to remedy this is, \eg, to choose smaller (and more) splats,
detect the problem and discard such splats, or introduce more complex texture
transformations. Since the proxy geometry in principle is to be short-lived, and
only a temporary geometry to be rendered before the server-rendered image can be
receieved, we have chosen a combination of the first two alternatives. Note that
implementing a more sophisticated texture coordinate transformation may actually
amount to performing the same work as for more and smaller splats. Equivalently
, using more and smaller splats may be regarded as a better texture transform
implementation.

%-------------------------------------------------------------------------
\subsubsection{Other splat considerations}

{\color{red} ... samle avsnittene nedenfor og komprimere...}

%-------------------------------------------------------------------------
\subsubsection{Splat sizing}

Note that as soon as a given proxy model is transformed, each splat should
ideally be rendered into a number of client pixels according to the new splats
3D position. A splat transformed into view closer to the observer, should result
in more client fragments, and vice versa. To achieve this, we simply use the
vectors $\sv_{\Delta x}$ and $\sv_{\Delta y}$ from the previous section.
In addition to this, we scale the splats up a bit, typically with a factor 2, so
that they overlap. The benefit of this is that we can render rectangular
window-aligned splats on the client, without getting large uncovered areas when
interactively rotating and scaling the model quickly on the client.

%-------------------------------------------------------------------------
\subsubsection{Splat depth fragments}

For larger splats, covering many fragments, it makes sense to also compute and
use depth fragments in the client's fragment shader. This will improve the usage
of more proxy models simultaneously, since the risk of a splat with a not very
representative depth value covers another and better splat, is reduced. The
``intra-splat'' depth values can easily be fetched from the depth image, just as
the texture is looked up for color, with the same texture transformation. One
problem is that many thin clients may not support this, in WebGL, the ``frag
depth'' feature is an extension, which may not be supported everywhere.

%------------------------------------------------------------------------
\subsection{Splat set replacement algorithms}
\label{sec:proxyModelReplacement}

In this paper we mainly concern ourselves with proxy models defined as sets of
splats. Since each splat set is the product of one instance of a server-rendered
image, depth buffer and viewing parameters, it makes sense to retain more than
one such model on the client. By doing this, the client may combine them in
order to provide renderings for viewing parameters very different from the last
one received.

As soon as the client has a set of such proxy models, it will also need an
algorithm for replacing old models with new ones.  One can imagine a plethora of
{\em splat set replacement algorithms}, we have tested three approaches that all
retain a constant number of proxy models. The first algorithm is to replace the
one with a viewing direction differing the most from the newly received
model. The second simply replaces the oldest one in store. The third replaces a
proxy model $k$ if the replacement results in the following objective function
being reduced,
\[
  \text{coverage} =
  \sum_{i=1}^n 
    \sum_{j\neq i}^n 
      ( \angle(\textbf{camdir}_i, \textbf{camdir}_j )^2,
\]
where $n$ is the number of proxy models available and the model to be replaced
is the one that minimizes this after replacement.
%\textbf{droppe fra her: }
%\ie, if $\text{coverage}_k < \text{coverage}$, where
%\[
%  \text{coverage}_k =
%  \min_j
%  \sum_{i=1, i\neq j}^n 
%    ( \angle(\textbf{camdir}_i, \textbf{new\ camdir}_j )^2.
%\]
%\textbf{til hit?}
Here, $\textbf{camdir}$ is the direction in which the camera was looking for the
generation of that particular proxy model.  There are some issues to consider
with respect to panning and zooming that we will not discuss here due to the
limited space available.

%-------------------------------------------------------------------------
\subsection{Dynamic/heuristic method selection, depth compression, auto-tuning}
\label{sec:autoTuning}

>>>>

\begin{abstract}
\color{black}
   We describe an algorithm and implementation of automatic proxy model
   generation for a client/server remote rendering setup. A {\em proxy model} is
   a light-weight version of a main 3D model that is cheaper to render. We
   automatically derive this from the main model. The client is typically a
   tablet, for which we only assume availability of WebGL and Javascript. The
   server makes use of OpenGL and a web server. When the rate of received
   server-rendered images deteriorate, the client renders the proxy model, which
   is computed from depth buffers bundled with rendered images from the
   server. This algorithm requires little modification of the application
   itself.
   % Husk: Skal vaere en blank linje nedenfor

@@

{\color{black} Hand in hand with increasingly powerful rendering engines comes
ever increasing requirements on computational accuracy, power efficiency, data
sizes, scaling, etc. Also enhanced by new cloud-based approaches etc., an effect
is that interactivity still is a difficult issue. We consider a client/server
model for 3D rendering, addressing latency, bandwidth and scaling problems in a
novel way.

We introduce a {\em proxy model}, defined as a temporary model to be shown and
manipulated locally on the client while waiting for the appropriate image from
the server. Producing such proxy models can be difficult for many kinds of 3D
data, like in our main case in which we have an oil reservoir viewer
(FRView,~\cite{cloudviz}) that renders large {\em corner point grids}, together
with faults, oil wells, and more. See Figure~\ref{fig:FRView} above for an
example of both a fully completed rendering, and automatically generated proxy
geometries rendered in Google Chrome.
%
Our solution is to pass depth information from the server along with ordinary
rendered images. From this, a rudimentary 3D model is built, and with the RGB
image as a texture, this model can be manipulated and rendered on the client
while waiting for the next update from the server. If the client does not change
the position or orientation of the model too much, this proxy model rendering
integrates seemlessly with the slower stream of server-rendered frames.
%
Even if bandwidth and latency is not a problem, it may be desirable to let a
server of limited capacity serve many simultaneous users, hence limiting the
effective server time available for each one. Suitable scaling may still be
achieved using our solution. This will be commercialized in the new SINTEF
project \mbox{ArrowHead}~\cite{arrowhead}.}

@@

Our approach has some similarities to {\em image based rendering} (IBR)
techniques, with the difference that an important IBR problem would be the
reconstruction of a depth map from images, while we have access to the full
depth map from a rendering pass.  Another way to use depth maps similar to what
we do, is for ``immersive streaming'', see \eg,~\cite{ibr}, where focus is on
depth map compression, an issue that we also consider. {\color{black}Another work
focused on similar streaming and geometry compression,
is~\cite{teler}. Also,~\cite{220764} contains some of the same ideas as our
work, with respect to image-based rendering accelleration.}
%
Common to many IBR-algorithms is also that of stitching together 3D or 2.5D
point clouds. We could also in principle do this for our 2.5D maps fetched from
the server, but it is not clear that the benefit would outweight the
cost.
%
We are not aware of anybody else using depth maps exactly for our purpose, most
research seems to be focused on the retrieval of the depth map from images,
while we use an existing depth map to distill and render temporary geometry.
%
{\color{black}Finally, we note that these 2.5D heightmaps are exactly what ``3D cameras''
(time-of-flight and other range image sensors) produce, but most authors
considering these are typically building more complex geometries before
visualization.} % , see~\eg,~\cite{}
\marginpar{\textbf{REF?!}}

@@

{\color{black} Since the depth buffer is a height map seen from the observer, it
does not contain information about occluded scene elements. Our approach assumes
that small transformations of this height map still will give good
approximations of the scene. In Figure~\ref{fig:2DheightmapRotated} below, a
sequence of three server-rendered images (thick lines) is shown, together with
intermediate client-rendered proxy models with different features that will be
discussed in Section~\ref{sec:client}.}

\begin{figure}[htb]
  \centering
  \subfigure[Low frame rate]{
    \input{fig3a.tex}
  }
  \subfigure[Initial proxy model]{
    \input{fig3b.tex}
  }
  \subfigure[Client-generated frame]{
    \input{fig3c.tex}
  }
  \subfigure[With texturing]{
    \input{fig3d.tex}
  }
  \subfigure[Screen-space-sized splats]{
    \input{fig3e.tex}
  }
  \subfigure[With intra-splat depths]{
    \input{fig3f.tex}
  }
  \caption{\label{fig:2DheightmapRotated}
{\color{black} 
Bird's view of 3D model (solid lines) and proxy model renderings.
}
%           The depth map and image sent from the server enables the client to
%           recover a 3D model approximating the one rendered on the server. This
%           client-side model is shown here with a small additional rotation, again
%           seen from above.
}
\end{figure}

%Note that the depth map from the server does not allow the client to recover all
%information in the server side model. For instance, the solid surface shown in
%Figure~\ref{fig:2Dheightmap} (\ie, the {\em topology}) cannot be deduced, hence
%we show the dots in Figure~\ref{fig:2DheightmapRotated}~(b). Still, there are
%several things we can do to improve the client's approximation of
%the server-side image, and these will now be described.

%-------------------------------------------------------------------------
\subsection{The server side}

{\color{black}
The server renders the 3D model into a framebuffer, and in the process generates
a depth image that we also send to the client.
Since this adds to the data being sent, it must be kept to a minimum. We have
found that reducing the spatial resolution of the depth image (for instance by a
factor of $1/16$) only degrades the proxy model imperceptibly.  We also encode
each depth value in the range $[0, 1]$ as a 16 bit fixed point number, and the
bundling of the depth image with the rgb image then imposes just a small data
overhead.
Further compression may bring this down even more, but the cost of
compression/decompression must also be
considered. In~\cite{DBLP:journals/tvcg/Lindstrom14} a suitable algorithm is
proposed that promises to be fast both for the compression and decompression
stages.
}

%-------------------------------------------------------------------------
\subsection{The client side}
\label{sec:client}

{\color{black}

When the client receieves an rgb and depth image, together with view
transformations, it builds a ``proxy model'' from this. This model can then be
transformed and rendered directly, or it can be combined with other proxy models
the client already has in store, see Section~\ref{sec:proxyModelReplacement}.

As indicated by Figure~\ref{fig:2DheightmapRotated}~(b), the received height map
does not allow us more than concluding where a set of points belong on the 3D
model, \ie, we have little topologic information. Since the information from the
depth map can only contain the foremost point along any ray from the observer,
it is said to be in 2.5D, as opposed to 3D.  The simplest thing the client can
do, is just to transform and render the set of 3D points with color sampled from
the server-rendered image, as is illustrated in
Figure~\ref{fig:2DheightmapRotated}~(b).

Rendering a 3D point for each depth fragment available may tax a thin client,
and it may be necessary to use a smaller number of primitives and instead render
each with a larger number of pixels on the client side, a {\em splat}.  A set of
such splats for a given depth image we call a {\em proxy model}. By rendering
splats, we get fewer ``false connections'' than if we connect 3D points and
reconstruct topology, but we risk getting more ``holes'' in our models.

We can render each splat as a fixed geometry, \eg, a disk or rectangle, with the
correspondig color from the image, see Figure~\ref{fig:2DheightmapRotated}~(c),
where we have introduced a transformation local to the client.  We will briefly
describe some improvements to this. Note that other possibilities include
building and maintaining a 3D occupancy mesh, computing a distance field from
which iso-surfaces can be extracted, etc.
}

@@

{\color{black}
Each splat produces many client pixels, necessitating an ``intra-splat''
fragment texturing for which we need a local texture coordinate transform.
A first approximation is for the client to assume that the corresponding part of
the server's model is planar in a region around the given point. If this is the
case, a local 2D texture transformation will provide a good approximation to the
intra-splat texturing to be performed on the client. Let $P_c$ and $P_s$ be
projection matrices on the client and server, respectively, and $M_c$ and $M_s$
corresponding view matrices. For the splat centered in $\vv_{i, j} = (x_j, y_i,
z_{i, j})$, to be centered on the client's canvas at screen coordinate $\pv_{i,
j} = P_c M_c M_s^{-1} P_s^{-1}\vv_{i, j} = U\vv_{i, j}$, the texture coordinate
transformation to be used is,
\[
  T =
  \begin{pmatrix}
    \frac{1}{n_x} & 0 \\
    0 & \frac{1}{n_y}
  \end{pmatrix}
  \Big( \sv_x \, \, \, \sv_y \Big)^{-1}
  \begin{pmatrix} 
    \frac{w}{n_x} & 0 \\
     0 & \frac{h}{n_y}
  \end{pmatrix}
   =
  A
  \Big( \sv_x \, \, \, \sv_y \Big)^{-1}
  B,
\]
where $A$ maps the ``client's splat region'' (in $[0, 1]^2$) to the
corresponding texture area, $(\sv_{x} \, \, \, \sv_{y})^{-1}$ maps the
``client's screen space splat area'' to $[0, 1]^2$, and $B$ provides a scaling
factor to fill the {\em viewport} of size $w \times h$ with $n_x \times n_y$
splats laid out in a grid, when $M_c=M_s$.
%
% After mult by proj matrix: clip coo, range of (x, y, z) = [-w, w]^3
% After persp division: ndc [-1, 1]^3
% After viewport transf: window coordinates
% (Range [vp[0] vp[1]] x [[vp[2] vp[3]] x [0, 1] ?)
%
We obtain $\sv_{x}$ and $\sv_{y}$ by evaluating proxy model positions followed by 
perspective division and transformation into window coordinates,
\[
  \pv =
  U { \small \begin{pmatrix} x \\ y \\ d \\ 1 \end{pmatrix} },
  \pv_{x+\Delta x} =
  U { \small \begin{pmatrix} x+\Delta x \\ y \\ d_{\Delta x} \\ 1 \end{pmatrix} }
  \text{and}\, \, \, 
  \pv_{y+\Delta y} =
  U { \small \begin{pmatrix} x \\ y+\Delta y \\ d_{\Delta y} \\ 1 \end{pmatrix} },
\]
where $d = 2D(x, y) - 1$, $d_{\Delta x} =
2D(x+\Delta x, y) - 1$, and $d_{\Delta y} = 2D(x,
y+\Delta y) - 1$ are depths sampled from the received depth buffer
$D$ and transformed to $[-1, 1]$. With a
$w$-component set to one, we have in effect done a perspective division, so that
we are in clip space and multiplication with $U$ is appropriate.
%
Note that $\Delta x$ and $\Delta y$ are not unique, we use
\[
  \Delta x = n_x / \text{width}(D) \text{\,\,\,\,and\,\,\,\,}
  \Delta y = n_y / \text{height}(D),
%  \label{eq:deltaChoice}
\]
but something larger may also be used. These are used for looking up depth image
samples, and the calculation of $\sv_x$ and $\sv_y$ is really a gradient
approximation, so we should not make them to large either.

This leaves us $\pv$, $\pv_{x+\Delta x}$ and $\pv_{y+\Delta y}$ in clip
coordinates, and from this we get
corresponding window coordinates $\sv$, $\sv_{x+\Delta x}$ and $\sv_{y+\Delta
y}$, from which we finally get splat spanning vectors
\[
  \sv_{\delta} =
  \sv_{x+\delta} - \sv =
    \frac{L}{2\delta} \left(
        \frac{\pv_{x+\delta}.xy}{\pv_{x+\delta}.w} -
        \frac{\pv.xy}{\pv.w}
    \right),
\]
with $\delta=\Delta x$ or $\delta=\Delta y$, $L$ is viewport size, either $w$ or
$h$, and we have used the ``shader notation'' for vector components.  The client
renders a large \texttt{glPoint} for each splat, with texture coordinates $(s,
t)^\prime$, and each fragment then looks up the server-rendered image at
position
\[
  \begin{pmatrix}
    s \\ t
  \end{pmatrix} +
  T 
  \begin{pmatrix}
    u \\ v
  \end{pmatrix}
  =
  \begin{pmatrix}
    s \\ t
  \end{pmatrix} +
  T 
  \begin{pmatrix}
    \texttt{glPointCoord}-\frac{1}{2}
  \end{pmatrix}
\]
where $(u, v)$ are ``intra-splat texture coordinates''.  When the assumption
that the geometry is locally planar does not hold, \eg, if the splats are very
large, or they originate from a curved or non-smooth part, this may look rather
odd, see Figure~\ref{fig:LargeSplatsOnCorners}.

\begin{figure}[htb]
  \centering
  \subfigure[Server-rendered]{
    \includegraphics[width=.3\linewidth]{splat1.jpg}
  }
  \subfigure[$P_c M_c = P_s M_s$]{
    \includegraphics[width=.3\linewidth]{splat2.jpg}
  }
  \subfigure[$P_c M_c \neq P_s M_s$]{
    \includegraphics[width=.3\linewidth]{splat3.jpg}
  }
  \caption{\label{fig:LargeSplatsOnCorners}
           Splats for which depths are sampled on a planar region look nice on
  exactly this region, but contorted elsewhere, notice the corner and edges of
  the cube.}
\end{figure}
%\vspace{-4ex}

When the geometry is not planar, $T$ produces the wrong result for parts of a
splat.  Two ways to remedy this could be to choose either more and smaller
splats, or introduce more complex texture transformations. Note that using a
more sophisticated texture coordinate transform may amount to performing the
same work as for more and simpler transforms. The latter may be regarded as
exactly a better ``global'' texture transform implementation.
}

%-------------------------------------------------------------------------
\subsubsection{Other splat considerations}
\label{sec:proxyModelReplacement}

{\color{black}

\textbf{Splat sizing} 
Each splat should be rendered into a number of client pixels according to the
new splats 3D position. To achieve this, we use the vectors $\sv_{\Delta x}$ and
$\sv_{\Delta y}$ from the previous section, and in addition, we scale the splats
up a bit so that they overlap. Hence, we can render rectangular window-aligned
splats with less risk of getting uncovered areas when interactively rotating and
scaling the model on the client.

\textbf{Splat depth fragments}
For larger splats it makes sense to also compute and use depth fragments on the
client. The ``intra-splat'' depth values can easily be fetched from the depth
image, just as the texture is looked up for color. Not all clients support this
at the time of writing, it is an extension in WebGL currently.

\textbf{Splat set replacement algorithms}
We mainly concern ourselves with proxy models defined as sets of splats, and it
makes sense to keep a set of such models on the client, then we may combine them
to cover larger ranges of client-side transformations. One can imagine a plethora of
{\em splat set replacement algorithms}, we have tested three approaches that all
retain a constant number of proxy models. The first is to replace the
one with a viewing direction differing the most from the newly received
model. The second simply replaces the oldest one in store. The third replaces a
proxy model $k$ if the replacement results in the following objective function
being reduced,
\begin{equation}
  \text{coverage} =
  \sum_{i=1}^n
    \sum_{j\neq i}^n 
      ( \angle(\textbf{camdir}_i, \textbf{camdir}_j )^2,
  \label{eq:coverage}
\end{equation}
where $n$ is the number of proxy models available, $\textbf{camdir}$ is the
direction in which the camera was looking when a particular model was generated,
and the model to be replaced is the one that minimizes~(\ref{eq:coverage}) after
replacement.

}

%-------------------------------------------------------------------------
\subsection{Auto-tuning}
\label{sec:autoTuning}

{\color{black}

It is important that the process of generating and sending proxy model data from
the server itself does not slow down transmission.  There are mainly three
sources of delay for server-rendered images to the client; high latency, low
bandwidth, and slow server-rendering.  In the first case, it seems prudent to
have a better proxy model on the client, that can be used for longer time and
for interactivity resulting in viewing parameters deviating more from the last
received server-rendered image. In the two other cases, it is important for the
proxy model generation/transmission to be quicker, both in order to get the
proxy model to the client and keep from delaying the server-image more than
necessary.

We have adopted an adaptive specification of proxy model data from the server
involving a more light-weight image (lossy JPG-compression with adaptive quality
control) while interaction is ongoing.  Also, reduced resolution of the depth
buffer sent from the server is used. Another possibility is to let the client
dynamically set the number of splats, number of proxy models, etc.

}
